---
title: "hw2 to Jeff from gx2144"
author: "Guangling Xu"
date: "9/30/2019"
output: github_document
---

## Import and Clean data
```{r,  message=FALSE}
library(tidyverse)
library(readxl)
library(ggplot2)
```

```{r}
p1data = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 1 ,range = "A2:N388" ) %>%
  janitor::clean_names() %>% 
  na.omit(p1data, cols = " dumpster") %>% 
  mutate(
    sports_balls = as.integer(round(sports_balls), digits = 0)
  )

p1data1 = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 4, range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  na.omit(p1data1) %>% 
  mutate(year = c(2018),
         month = as.numeric(month),
         month = month.name[month]
   )

  
p1data2 = 
  read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 5 , range = "A2:B15") %>%
  janitor::clean_names() %>%
  na.omit(p1data2) %>% 
  mutate(year = c(2017),
         month = as.numeric(month),
         month = month.name[month]
  )

mergedata = full_join(p1data1 , p1data2) 
mergedata
```

## Summary data

* There are 337 rows and 14 columns in sheet"Mr.Trash Wheel".
 Vriable names are "dumpster","month","year","date","weight (tons)","volume (cubic yards)",	"plastic bottles","polystyrene","cigarette butts","glass bottles","grocery bags","chip bags",	"sports balls","homes powered*"
 
 "dumpster" means the number of bins;
 
 "month", "day", "year" are varaibles denoting the date of collection
 
 "weight" and "volumn" are variables denoting the amount of the trash
 
 "plastic bottles","polystyrene","cigarette butts","glass bottles","grocery bags","chip bags"are variables denoting the type of the trash
 
* There are 18 rows and 3 columns in sheet"mergedata".
 Variable names are "month","total" and "year"
 
* The total precipitation in 2018 is 23.50

* The median number of sports balls in a dumpster in 2017 is 10

## Problem 2
## Import and Clean Data
```{r}
pols_month = 
  read.csv("./data/pols-month.csv") %>% 
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month","day"), sep = "-") %>%
  mutate(
         month = as.numeric(month),
         month = month.abb[month],
         president = recode(prez_gop, "0" = "gop", "1" = "dem" , "2" = "gop2")
         ) %>% 
  select(-prez_dem) %>% 
  select(-prez_gop) %>% 
  select(-day)
         

snp = 
  read.csv("./data/snp.csv") %>% 
  janitor::clean_names() %>%
  separate(date, into = c("month", "day", "year"), sep = "/") %>%
  mutate( 
    month = as.numeric(month)
  ) %>% 
  arrange(year,month) %>% 
  select("year","month",everything()) %>% 
  mutate(
    month = month.abb[month])

unemplyment = 
  read.csv("./data/unemployment.csv")
unemplyment = 
  pivot_longer(
    unemplyment, 
    Jan:Dec,
    names_to = "month", 
    values_to = "unemplyment") %>% 
  rename(year = Year)  %>%
  mutate(
    year = as.character(year)
  )
  

```

## Merge data
```{r}
result = 
  left_join(pols_month,snp)
finalresult = 
  left_join(result, unemplyment, by = "year")
head(finalresult,5)
tail(finalresult,5)## show part of the table
```


## Problem 3
## Import and Clean Data
```{r}
names = 
  read.csv("./data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(
    gender = str_to_lower(gender),
    ethnicity = str_to_lower(ethnicity),
    ethnicity = recode(ethnicity, "asian and paci" = "asian and pacific islander",
                       "black non hisp" = "black non hispanic" , 
                       "white non hisp" = "white non hispanic"),
    
   child_s_first_name = str_to_title(child_s_first_name)
   ) %>% 
  distinct()
```


## Filter data of Olivia
```{r}
olivia = 
  filter(names, child_s_first_name == "Olivia" , gender == "female") %>% 
  select(-count) 
pivot_wider(olivia, 
              names_from = year_of_birth, 
              values_from = rank)  

```

## Scatterplot
```{r}
male_name = 
    filter(names, gender == "male" ,year_of_birth == "2016", 
           ethnicity == "white non hispanic" ) 
male_nameplot = ggplot(male_name, aes(x = rank, y = count ))+
  geom_point()+
  geom_smooth(se = FALSE)
male_nameplot
ggsave("male_name.pdf" , male_nameplot , width = 8, height = 5 )
```


